---
title: "Project01"
author: "Li Du"
date: "October 5, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare the analysis

```{r loadLibs}
# Remember to install packages if they are not already installed
library(corrplot)
library(MASS)
library(GGally)
library(ggplot2)
```

A simple function was defined to classify and print out the level of correlation
```{r classify}
classify <- function(x) {
  if (abs(x) >= 0.7) { lev = "strongly"}
  else if (abs(x) >= 0.4 & abs(x) < 0.7) {lev = "moderately"}
  else lev = "poorly"
  
  if (x > 0) {s = "positively"}
  else s = "negatively"
  sprintf("The median value of owner-occupied house and pupil-teacher ratio by town is %s and %s correlated", lev, s)
}
```

A function was created to interpret the result of ANOVA test
```{r ANOVAResult}
ANOVARes <- function (x) {
  alpha = 0.05 #significance level = 5%
  if (x[[1]][["Pr(>F)"]][[1]] >= alpha) {
    print("The median house values and the pupil-teacher ratios are independent")
  }
  else 
    print("The median house values and the pupil-teacher ratios are NOT independent")
}
```


Import the "Boston" dataset and perform brief cleaning
```{r dataset}
BosHousing <- Boston
BosHousing_num <- BosHousing[,!(names(BosHousing) %in% "chas")] #Remove the variable "chas" since it is a boolean-type variable
```


## Explore and answer the SMART question 

### correlation
```{r correlation}
correlation = cor(BosHousing_num$medv, BosHousing_num$ptratio)
sprintf("The correlation coefficient is %f", correlation)
classify(correlation)
```

### Scatter plot
```{r scatterplot, echo=FALSE}
ggplot(data=BosHousing, aes(x=ptratio, y=medv))+
  geom_point(shape=1)+ 
  geom_smooth(method=lm, se=FALSE) +
               labs(y="Median House Value (in $1000)", x="Pupil-teacher ratio by town")
```

## One-way ANOVA test

The dataset was divided into three parts based on the tertiles of "ptratio". Then, the median values of houses are extracted for the three groups. This new dataset MedV.aov will be used for the following ANOVA analysis.
```{r Pre-ANOVA}
nr <- nrow(BosHousing_num) #calculate the number of rows
BosHousing_num <- BosHousing_num[order(BosHousing_num$ptratio),]# order the data frame by "ptratio"
ptratio_lv <- c(rep("low",round(nr/3,0)),rep("mid",round(nr/3,0)),rep("hi",round((nr/3-1),0)))
MedV.aov <- data.frame(BosHousing_num$ptratio,ptratio)
colnames(MedV.aov) <- c("MedV", "ptratio")
```

Perform one-way ANOVA test
```{r ANOVA}
results <- aov(MedV~ptratio, data=MedV.aov)
summary(results)

ANOVARes(summary(results))
```


## Exploratory or general analysis of the whole datast

We can discuss whether this is needed and where to fit them
```{r TBD}
#exploratory analysis
corrplot(cor(BosHousing_num), type = "upper", is.corr = TRUE)
ggpairs(BosHousing_num)

#Whole dataset exploration


summary(BosHousing$ptratio)
summary(BosHousing$medv)

#Outliers

str(BosHousing) #All variables are integers or numeric

colSums(is.na(BosHousing)) #No missing values

# Applying Freedman/Diaconis rule p.120 ("Algorithm that chooses bin widths and locations automatically, based on the sample size and the spread of the data" 

#Write function for plots, save each in directoy

Bos_plots <- function(x,y) {
  hist(x, breaks = "FD", col = y) #Add chart title
  boxplot(x, col = "blue")
}


par(mfrow = c(1,2), lwd = 2)
dotchart(BosHousing$ptratio, lcolor = 0)
dotchart(sort(BosHousing$ptratio), lcolor = 0)

# Do males and females have similar variance in Head size?
boxplot(BosHousing$ptratio ~ BosHousing$medv)

with(BosHousing, hist(BosHousing$ptratio, breaks="FD", col="green"))
with(BosHousing, hist(BosHousing$medv, breaks="FD", col="blue"))
with(BosHousing, boxplot(BosHousing$medv, breaks="FD", col="blue"))

plot(BosHousing$medv, BosHousing$ptratio) #plot median house value against pupil-ratio

scatterplot(ptratio~medv, id.method="identify", data=BosHousing)

boxplot(BosHousing$ptratio)

boxplot(BosHousing$medv)

plot(BosHousing$medv)

#Chi-square

#Cramer's V

#Fisher's exact

```




